\textbf{\textit{Публікації в наукових журналах:}}
\begin{enumerate}[leftmargin=0.8cm]
	\item
	Sushko M. Ya. Conductivity and permittivity of dispersed systems 
	with penetrable particle-host interphase~/ M.~Ya.~Sushko, 
	A.~K.~Semenov~// Cond. Matter Phys. --- 2013 --- Vol.~16 --- No. 1 
	--- P.~13401.
	
	\item
	Semenov~A.~K. On applicability of differential mixing rules for
	statistically homo\-geneous and isotropic dispersions~/ A.~K.~Semenov~//
	J. Phys. Commun. --- 2018. --- Vol.~2. --- No. 3 --- P.~035045.
	
	\item
	Sushko~M.~Ya. A mesoscopic model for the effective electrical 
	conductivity of composite polymeric electrolytes.~/ M.~Ya.~Sushko,
	A.~K.~Se\-menov~// J. Mol. Liq. --- 2019. --- Vol.~279 --- P.~677.
	
	\item
	Sushko~M.~Ya. Rigorously solvable model for the electrical conductivity of dispersions of hard-core--penetrable-shell particles and its applications~/
	M.~Ya.~Su\-shko, A.~K.~Semenov~//
	Phys. Rev. E --- 2019. --- Vol.~100. --- P.~052601.
\end{enumerate}


\vskip 15pt
\textbf{\textit{Тези доповідей на наукових конференціях:}}
\begin{enumerate}[leftmargin=0.8cm, resume]
	\item Semenov A. Complex permittivity of disperse systems with penetrable particle-host interphase / A. Semenov, M. Sushko // 4-th International Conference on Statistical Physics: Modern Trends and Applications, 3-6 Jul., 2012: abstract - Lviv (Ukraine), 2012.
	
	\item Семенов А.К. Роль межфазной границы в формировании проводимости  и диэлектрической проницаемости мелкодисперсных систем / А.К. Семенов, М.Я. Сушко // 25-th International Conference: Disperse Systems, 17-21 Sept., 2012: abstract - Odesa (Ukraine), 2012.
	
	\item Sushko M. Ya. Finding the parameters of the interphase layers in fine dispersions with dielectric spectroscopy studies near the electrical percolation threshold / M.Ya. Sushko, A.K. Semenov // 5-th International Symposium: Methods and Applications of Computational Chemistry, 1-5 Jul., 2013: abstract - Kharkiv (Ukraine), 2013.
	
	\item Sushko M. Ya. Effect of interphase on the effective electrophysical para\-meters of fine dispersions and nanofluids / M.Ya. Sushko, A.K. Semenov // 6-th International Conference Physics  of  Liquid  Matter:  Modern Prob\-lems, 23-27 May, 2014: abstract - Kyiv (Ukraine), 2014.
	
	\item Gotsulskiy V. Ya. Conductivity and permittivity of isopropyl alcohol- and water-based nanofluids: experimental results, theory and inferences / V.Ya. Gotsulskiy, A.K. Semenov, M.V. Stiranec, and M.Ya. Sushko // 6-th International Conference Physics  of  Liquid  Matter:  Modern Prob\-lems, 23-27 May, 2014: abstract - Kyiv (Ukraine), 2014.
	
	\item Семенов А. К. Диэлектрическая проницаемость и проводимость дисперсных систем с неоднородной межфазной границей / А.К. Семенов, М.Я. Сушко // 26-th International Conference: Disperse Systems, 22-26 Sept., 2014: abstract - Odesa (Ukraine), 2014.
	
	\item Semenov A. K. A model for conductivity and permittivity of heterogeneous systems with complex microstructures / A.K. Semenov, M.Ya. Sushko // 2015 International Young Scientists Forum on Applied Physics, 29 Sept.-2 Oct., 2015: abstract - Dnipropetrovsk (Ukraine), 2015.
	
	\item Бабий К. А Особенности электрической проводимости дисперсных систем на основе полимерных матриц / К.А. Бабий, А.К. Семенов, М.Я. Сушко // 27-th International Conference: Disperse Systems, 19-23 Sept., 2016: abstract - Odesa (Ukraine), 2016.
	
	\item Семенов А. К. Роль міжфазних шарів у формуванні провідних та діелектричних властивостей дісперсноподібних систем: модель та застосування / А.К. Семенов, М.Я. Сушко // International conference: Development of innovation in the technical, physical and mathematical fields of sciences, 22-24 Sept., 2016: abstract - Mykolayiv (Ukraine), 2016.
	
%	\item Semenov A. K. Is the classical differential scheme for permittivity of emulsions consistent? / A.K. Semenov // IX Conference of Young Scientists ``Problems of Theoretical Physics'', 04-05 Dec., 2018: abstract - Kyiv (Ukraine), 2018.
	
	\item Sushko M. Ya. Effective electrical conductivity of composite polymer elec\-trolytes / M.Ya. Sushko, A.K. Semenov // 8-th International  Conference Physics  of  Liquid  Matter: Modern Problems, 18-22 May, 2018: abstract - Kyiv (Ukraine), 2018.
	
	\item Semenov A. K. Is the classical differential scheme for permittivity of emulsions consistent? / A.K. Semenov // 8-th International  Conference Physics  of  Liquid  Matter: Modern Problems, 18-22 May, 2018: abstract - Kyiv (Ukraine), 2018.
	
	\item Sushko M. Ya. Recent developments in the theory of electrodynamic homogenization of random particulate systems / M.Ya. Sushko, A.K. Semenov // 5-th International Conference on Statistical Physics: Modern Trends and Applications, 3-6 Jul., 2019: abstract - Lviv (Ukraine), 2019.
	
	\item Semenov A. K. Hard-core–penetrable-shell model for effective electric parameters of random particulate systems / A.K. Semenov, M.Ya. Sushko // 7-th International Conference: Nanotechnologies and Nanomaterials, 27-30 Aug., 2019: abstract - Lviv (Ukraine), 2019.
	
	\item Семенов А. К. Моделювання електрофізичного відгуку дисперсних систем з твердим дисперсійним середовищем / А.К. Семенов, М.Я. Сушко // 28-th International Conference: Disperse Systems, 16-20 Sept., 2019: abstract - Odesa (Ukraine), 2019.
\end{enumerate}